---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data  

The packages required to run and reproduce the code are listed.  

1. dplyr
2. ggplot2
3. gridExtra
4. taRifx
5. lattice

```{r library, echo=FALSE, message=FALSE}

library(dplyr)
library(ggplot2)
library(gridExtra)
library(taRifx)
library(lattice)
```

This code supose that file "activity.zip" is available at working directory.  

```{r load_prepare_data}

unzip("activity.zip")
activity <-  read.csv("activity.csv", sep = ",") %>%
  transform(steps <- as.numeric(steps)) %>%
  remove.factors() #Factors where removed to avoid be plotted on graphics
activity$date <- as.Date(activity$date, "%Y-%m-%d")

```

## What is mean total number of steps taken per day?

For this part of task it is used only the complete cases of data set.  

```{r daily_activity_complete_cases}

complete_activity <- complete.cases(activity)
act_no_NA <- activity[complete_activity,]

# Using aggregate funcion grouping data by day and calculating sum, mean and median

sum_steps_day <- aggregate(act_no_NA$steps, list(act_no_NA$date), sum)
mean_steps_day <- aggregate(act_no_NA$steps, list(act_no_NA$date), mean)
median_steps_day <- aggregate(act_no_NA$steps, list(act_no_NA$date), median) #The value is 0 for each day
```

The mean and median for the total of steps per day are respectively:

```{r mean_med}
total_mean_steps_day <-  mean(sum_steps_day$x)
total_median_steps_day <- median(sum_steps_day$x)
print(paste("Mean of total of steps a day :", total_mean_steps_day))
print(paste("Median of total of steps a day :", total_median_steps_day))
```

Considering this values are so close, only the median value is shown in the histogram.

```{r histogram}
hist(sum_steps_day$x, breaks = 20, xlab = "Number of Steps per Day", 
     main = "Histogram for number of steps a day")
abline(v=total_median_steps_day,col="blue")
text(total_median_steps_day+350,9,"Median", col="blue", adj=c(0,0.5))
```




## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
